---
title: "IST707GP"
author: "Fu, Maloney, Tran, Ventosa"
date: "7/26/2020"
output:
  word_document: default
---

Introduction:

Auto insurance is mandated in most states to provide drivers and automobile owners financial protection in the event of an accident. Laws vary from state to state and insurance coverages are designed to meet state laws while also providing customized policies to meet individual needs.  Insurance policies are generally issued for six-month to one-year timeframes. When nearing a policy's expiration, insurance companies send renewal offers to retain customers and collect premiums. As a highly competitive industry customer churn rate averages at approximately 12% every year (https://bit.ly/3fgvVXv) making customer retention one of the most important factors for success. 

FMTV Auto Insurance maintains and updates its customer records in order to discover new and effective ways to retain its clientele. The company provides both personal and company auto insurance and attempts to offer customized insurance renewals to best meet their clients' varying needs. As part of its marketing efforts to retain clients, FMTV sends out one of four renewal offers to customers nearing the end of their policy terms. A purchase or non-purchase response to the offers are recorded to determine the marketing return on investment. Recently purchase response rates to their renewal offers have been down. The marketing department questions whether a more effective method of distributing renewal offers may increase purchases and reduce customer churn. 

Customer Lifetime Value (CLV) is one ot the key metrics FMTV uses to measure its success. CLV is a measures of a customer's total worth to a business over the entire time period of the customer-business relationship. By developing strategies to reduce client churn, policy term businesses can increase the overall CLV of its clients.  Higher retention also leads to lower costs and better profit margins as retaining existing customers costs less than acquiring new ones. However, FMTV has kept track of its clients' CLV not only as a measure of success but to also develop predictive models to identify potentially high CLV customers to acquire or maintain.

The simplest formula for measuring CLV is:

Customer revenue per year 
multiplied by 
Duration of the relationship in years 
minus 
Total costs of acquiring and serving the customer 
equals to 
CLV

Analysis and Models:

About the Data:

Loading Data
```{r}
auto <-read.csv("https://raw.githubusercontent.com/jcventosa/IST-707/master/WA_Fn-UseC_-Marketing-Customer-Value-Analysis.csv")
```

Data Structure
```{r structure}
str(auto)
```
The data has 9134 records with the following 24 variables:

Variables                       Description
Customer:                       Customer ID number
State:                          State of residence or business
Customer Lifetime Value:        Customer's total worth to business over life of the relationship
Response:                       Yes or No response to a renewal offer
Coverage:                       Type of policy (Basic, Extended, Premium)
Education:                      Level of education (High School or less, College, BA, MA , PHD)
Effective To Date:              Date the policy expires
Employment Status:              (Employed, Unemployed, Retired, Disabled, Medical Leave)
Gender:                         Male or Female
Income:                         Customers' annual income
Location Code:                  (Rural, Suburban, Urban)
Marital Status:                 (Single, Married, Divorced)
Monthly Premium Auto            Amount of customers' monthly insurance payments
Months Since Last Claim         Number of months between customers' last reported insurance claim
Months Since Policy Inception   Number of months since customer began an insurance policy
Number of Open Complaints       Number of unresolved customer complaints
Number of Policies              Number of policies customer currently owns
Policy Type                     (Corporate Auto, Personal Auto, Special Auto)
Policy                          3 levels (L1, L2, L3) per Policy Type (Corporate, Personal, Special)
Renew Offer Type                4 types of renewal offers (Offer 1, Offer 2, Offer 3, Offer 4)
Sales Channel                   Channels to purchase a policy (Agent, Branch, Call Center, Web)
Total Claim Amount              Cummulative amount of claims since policy inception
Vehicle Class                   Type of vehicle (4-Door, Luxury, Luxury SUV, Sports Car, SUV, 2-Door)
Vehicle Size                    (Large, Medium, Small)

Data Cleaning:

Check for NA, missing and duplicated data
```{r echo=TRUE, message=FALSE, warning=FALSE}
#install.packages("rapportools")
library(rapportools)
sum(is.na(auto))
nrow(auto[duplicated(auto),])
```
No missing or duplicated data.

Check data summary for unusual distributions or incorrect data types
```{r summary, echo=TRUE, message=FALSE, warning=FALSE}
summary(auto)
```
Initial inspection of the dataset summary provides the proportional split for some key nominal variables as follows:

Response:         0.14 Yes        0.86  No
Gender:           0.51 Female     0.49  Male
Marital Status    0.58 Married    0.15  Divorced    0.27 Single
Location Code     0.19 Rural      0.63  Suburban    0.17 Urban
Coverage          0.61 Basic      0.30  Extended    0.09 Premium
Policy Type       0.74 Personal   0.04  Special     0.22 Corporate
Vehicle Size      0.10 Large      0.70  Medsize     0.19 Small

Data Transformation:

Converting the Effective.To.Date variable's data type from factor to date.  
```{r echo=TRUE, message=FALSE, warning=FALSE}
auto$Effective.To.Date <- as.Date(auto$Effective.To.Date, format = "%m/%d/%y")
```

In addition the following discretized variables will be added to the dataset.

CLV_Bins        Discretized Customer.Lifetime.Value 
IncomeBin       Discretized Income
MoPremiumBin    Discretized Monthly.Premium
ClaimBin        Discretized Total.Claim

```{r Discretization, echo=TRUE, message=FALSE, warning=FALSE}
# Adding columns for CLV, Income, Monthly Premium and Total Claim bins.

# Cutomer Lifetime Value bins
auto$CLV_Bins <- cut(auto$Customer.Lifetime.Value, 
                       breaks = c(0,1999,3999,5999,7999,9999,Inf),
                       labels = c("< $2000", "$2000-$3999","$4000-$5999",
                                  "$6000-$7999", "$8000-$9999", "$10000+"))
# Income bins
auto$IncomeBin <- cut(auto$Income, breaks = c(0,14999,29999,44999,59999,74999,Inf),
                    labels = c("< $15000", "$15000-$29999","$30000-$44999",
                               "$45000-$59999", "$60000-$74999", "$75000+"))
# Monthly Premium bins
auto$MoPremiumBin <- cut(auto$Monthly.Premium.Auto, 
                            breaks = c(0,74,99,124,149,Inf),
                            labels = c("< $75", "$75-$99","$100-$124","$125-$149","$150+"))

# Total Claim bins
auto$ClaimBin <- cut(auto$Total.Claim.Amount, 
                        breaks = c(0,249,499,749,999,Inf),
                        labels = c("< $250", "$250-$499","$500-$749","$750-$999","$1000+"))

# Spot check bin labels match appropriate values
head(auto[,c(3,25,10,26,13,27,22,28)],30)

# Replace NA in IncomeBin with "< $15000"
auto$IncomeBin[is.na(auto$IncomeBin)] <- "< $15000"
sum(is.na(auto$IncomeBin))
head(auto[,c(10,26)],15)
```

```{r eval=FALSE, include=FALSE}
# Write the cleaned data to a csv file named auto.csv
write.csv(auto,"~/Desktop/auto.csv")
```
Exploratory Data Analysis:

```{r}
hist(auto$Income)
hist(auto$Total.Claim.Amount)
hist(auto$Monthly.Premium.Auto)
hist(auto$Months.Since.Last.Claim)
hist(auto$Months.Since.Policy.Inception)
hist(auto$Customer.Lifetime.Value)
```

Client demographic background by Policy Type
```{r}
#install.packages("ggplot2", "RColorBrewer", "gridExtra"")
library(ggplot2)
library(RColorBrewer)
library(gridExtra)

PTGender <- ggplot(auto, aes(x= Policy.Type, fill=Gender)) + geom_bar(stat = "count") + 
  xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Gender") +
  scale_fill_brewer (palette = "Set1", labels = c("Female", "Male"))

PTMarital <- ggplot(auto, aes(x=Policy.Type, fill = Marital.Status)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Marital Status") +
  scale_fill_brewer (palette = "Dark2", name = "Marital Status") 

table(auto$Policy.Type, auto$Gender)
table(auto$Policy.Type, auto$Marital.Status)
grid.arrange(PTGender, PTMarital)
```
There is an approximate 50/50 split by Gender for both Corporate and Personal policy types with females just sligthly outnumbering men by a percent or less.  Women make up 55% of the Special auto policy types. The distribution of policy types by marital status is within +/- 4% of the overall distribution of 15% Divorced, 58% Married and 27% Single.

```{r}
PTState <- ggplot(auto, aes(x=Policy.Type, fill = State)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by State") +
  scale_fill_brewer (palette = "Accent", name = "State") 

PTLocation <- ggplot(auto, aes(x=Policy.Type, fill = Location.Code)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Region") +
  scale_fill_brewer (palette = "Dark2", name = "Region") 

table(auto$Policy.Type, auto$State)
table(auto$Policy.Type, auto$Location.Code)
grid.arrange(PTState, PTLocation) 
```
The majority of all three policy types are in California followed by Oregon, Arizona, Nevada and Washington. Suburban regions have over three times the number of Corporate and Personal policy types compared to Rural and Urban regions. Sixty-0ne percent of Special policy types are also located in Surban regions with Rural and Urban each making up approximately 14%-15%.  

```{r}
auto$Education <- factor(auto$Education, levels = c("High School or Below", "College", 
                                                    "Bachelor", "Master", "Doctor"))
auto$EmploymentStatus <- factor(auto$EmploymentStatus, levels = c("Employed", "Unemployed",
                                                        "Retired", "Disabled", "Medical Leave"))

PTEmployment <- ggplot(auto, aes(x=Policy.Type, fill = EmploymentStatus)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Employment Status") +
  scale_fill_brewer (palette = "Set1", name = "Employment Status")

PTIncomeBin <- ggplot(auto, aes(x=Policy.Type, fill = IncomeBin)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Income") +
  scale_fill_brewer (palette = "Accent", name = "Income") 

PTEducation <- ggplot(auto, aes(x=Policy.Type, fill = Education)) + geom_bar(stat = "count") +
    xlab("Policy Type") + ylab("Policy Count") + ggtitle("Policy Type by Education Level") +
  scale_fill_brewer (palette = "Accent", name = "Education") 

table(auto$Policy.Type, auto$Education)
table(auto$Policy.Type, auto$EmploymentStatus)
table(auto$Policy.Type, auto$IncomeBin)
PTEducation
PTEmployment
PTIncomeBin
```
There is nearly an equal distribution of clients with High School or Less, College and Bachelors where each education level makes up approximately 30% of each policy type.  The synchronicity in Employed and Unemployed is also true making up approximately 62% and 25% of each policy type respectively. Lastly 44%-45% all three policy types have incomes less than $30,000.

Exploring Renewal Offers

```{r}
table(auto$Renew.Offer.Type, auto$Policy.Type)

# MyTheme to format font sizes for my graphs
MyTheme <- theme(axis.text = element_text(size = 14), 
          axis.title = element_text(size = 16, face ="bold"), 
          plot.title = element_text(size = 20, face = "bold"))
MyTheme2 <- theme(axis.text = element_text(size = 10), 
                  axis.title = element_text(size = 18, face ="bold"), 
                  plot.title = element_text(size = 20, face = "bold"), 
                  axis.text.x = element_text(angle = 90))

ggplot(auto, aes(x=Renew.Offer.Type, fill = Policy.Type)) + geom_bar(stat="count")+
  xlab("Renewal Offer Type") + ylab("Policy Count") + ggtitle("Renewal Offer Type by Policy Type") +
  scale_fill_brewer(palette = "YlOrBr") + MyTheme

ggplot(auto, aes(x=Renew.Offer.Type, fill = Policy.Type)) + geom_bar(stat="count", position = "fill")+
  xlab("Renewal Offer Type") + ylab("") + ggtitle("Percent of Renewal Offer Types by Policy Type") +
  scale_fill_brewer(palette = "YlOrBr") + MyTheme
```
Each client from the three policy types received a single renewal offer type of 1, 2, 3 or 4.  Offer 1 had the highest distribution; received by 41% of the clients.  Offer 2 was received by 32%, Offer 3 by 16% and Offer 4 received by 11% of the clients. Interestingly the offer types were distributed at nearly equal percentages by each policy type.

```{r}
ROGender <- ggplot(auto, aes(x= Renew.Offer.Type, fill=Gender)) + 
  geom_bar(stat = "count", position = "fill") + 
  xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Type by Gender") +
  scale_fill_brewer (palette = "Set1", labels = c("Female", "Male"))

ROMarital <- ggplot(auto, aes(x=Renew.Offer.Type, fill = Marital.Status)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Type by Marital Status") +
  scale_fill_brewer (palette = "Dark2", name = "Marital Status") 

table(auto$Renew.Offer.Type, auto$Gender)
table(auto$Renew.Offer.Type, auto$Marital.Status)
grid.arrange(ROGender, ROMarital)
```
The offers were also distributed at nearly equal proportions by Gender and Marital Status

```{r}

ROState <- ggplot(auto, aes(x=Renew.Offer.Type, fill = State)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Types by State") +
  scale_fill_brewer (palette = "Accent", name = "State") 

ROLocation <- ggplot(auto, aes(x=Renew.Offer.Type, fill = Location.Code)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Types by Region") +
  scale_fill_brewer (palette = "Dark2", name = "Region") 

table(auto$Renew.Offer.Type, auto$State)
table(auto$Renew.Offer.Type, auto$Location.Code)
grid.arrange(ROState, ROLocation) 
```
The Renewal Offers were also distributed at nearly equal proportions by State and Region

```{r}
Mytheme <- theme(axis.text = element_text(size = 12), 
          axis.title = element_text(size = 14, face ="bold"), 
          plot.title = element_text(size = 16, face = "bold"))

ROEmployment <- ggplot(auto, aes(x=Renew.Offer.Type, fill = EmploymentStatus)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Type by Employment Status") +
  scale_fill_brewer (palette = "YlOrBr", name = "Employment Status") + Mytheme

ROIncomeBin <- ggplot(auto, aes(x=Renew.Offer.Type, fill = IncomeBin)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Type by Income") +
  scale_fill_brewer (palette = "OrRd", name = "Income") + Mytheme

ROEducation <- ggplot(auto, aes(x=Renew.Offer.Type, fill = Education)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Renew Offer Type") + ylab("") + 
  ggtitle("Percent of Renew Offer Type by Education Level") +
  scale_fill_brewer (palette = "Oranges", name = "Education") + Mytheme

table(auto$Renew.Offer.Type, auto$Education)
table(auto$Renew.Offer.Type, auto$EmploymentStatus)
table(auto$Renew.Offer.Type, auto$IncomeBin)
ROEducation
ROEmployment
ROIncomeBin
```
There are minor differences in how the offers were distributed by education level with Offer 1 more heavily distributed among College and Bachelor than other offers. Offer 1 was also received by a lower proportion of employed clients compared to the other offers. This difference was also evident in the income levels as nearly 70% of Offer 1 recipients had incomes below $45k compared to 51% to 59% of recipients of Offer 2, 3 and 4. 

```{r}
library(ggplot2)
library(RColorBrewer)
ggplot(auto, aes(x=Renew.Offer.Type, fill = Response)) + geom_bar(stat = "count", position = "fill") +
  xlab("Renew Offer Type") + ylab("") + ggtitle("Percent of Renew Offer Type by Response") +
  scale_fill_manual(values = c("darkorange1", "darkorange4")) + theme(axis.text = element_text(size = 12), 
          axis.title = element_text(size = 14, face ="bold"), 
          plot.title = element_text(size = 18, face = "bold"))
```
The purchase response rates for the renew offers were very low. Offer 2 had the best purchase response rate at 23.4% followed by Offer 1 with a purchase response rate of 15.8%. Only 2% of Offer 3 recipients made a renewal while Offer 4 was declined by all client recipients.

```{r}
RPSGender <- ggplot(auto, aes(x= Gender, fill=Response)) + 
  geom_bar(stat = "count", position = "fill") + 
  xlab("Gender") + ylab("") + 
  ggtitle("Percent of Gender by Response") +
  scale_fill_brewer (palette = "Set1")

RPSMarital <- ggplot(auto, aes(x=Marital.Status, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Marital.Status") + ylab("") + 
  ggtitle("Percent of Marital Status by Response") +
  scale_fill_brewer (palette = "Dark2") 

table(auto$Response, auto$Gender)
table(auto$Response, auto$Marital.Status)
grid.arrange(RPSGender, RPSMarital)
```
Men and women were equally as likely to respond with a "Yes".  Divorced clients were slightly more likely to say "Yes" to an offer.

```{r}

RPSState <- ggplot(auto, aes(x=State, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("State") + ylab("") + 
  ggtitle("Percent of State Clients by Response") +
  scale_fill_brewer (palette = "Accent") 

RPSLocation <- ggplot(auto, aes(x=Location.Code, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Region") + ylab("") + 
  ggtitle("Percent of Region by Response") +
  scale_fill_brewer (palette = "Dark2") 

table(auto$Response, auto$State)
table(auto$Response, auto$Location.Code)
grid.arrange(RPSState, RPSLocation) 
```
The purchase response rates were fairly the same by state.  However, clients from Suburban areas did have a higher purchase response rate than Rural or Urban clients.

```{r}
RPSEmployment <- ggplot(auto, aes(x=EmploymentStatus, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Employment Status") + ylab("") + 
  ggtitle("Response Rate by Employment Status") +
  scale_fill_manual (values = c("darkorange1", "darkorange4")) + Mytheme

RPSIncomeBin <- ggplot(auto, aes(x=IncomeBin, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Annual Income") + ylab("") + 
  ggtitle("Response Rate by Income") +
  scale_fill_brewer (palette = "Accent") + MyTheme2

RPSEducation <- ggplot(auto, aes(x=Education, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Education Level") + ylab("") + 
  ggtitle("Response Rate by Education Level") +
  scale_fill_brewer (palette = "YlOrBr", name = "Education") + MyTheme2

table(auto$Response, auto$Education)
table(auto$Response, auto$EmploymentStatus)
table(auto$Response, auto$IncomeBin)
RPSEducation
RPSEmployment
RPSIncomeBin
```
Clients with post graduate degrees were slightly more likely to accept the renewal offers they received. Nearly 75% or retired clients responded "Yes" to their renewal offers.  Clients earning between 15,000-29,999 and 60,000-74,999 answered "Yes" to renewal offers more often than clients of other income ranges.

```{r}
Mytheme2 <- theme(axis.text = element_text(size = 10), 
                  axis.title = element_text(size = 14, face ="bold"), 
                  plot.title = element_text(size = 16, face = "bold"), 
                  axis.text.x = element_text(angle = 90))

ggplot(auto, aes(x=CLV_Bins, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Customer Lifetime Value") + ylab("Percent") + 
  ggtitle("Response Rate by Customer Lifetime Value") +
  scale_fill_brewer (palette = "Dark2", name = "Response") + Mytheme2
```

```{r}
Offer1 <- auto[auto$Renew.Offer.Type=="Offer1",]
Offer2 <- auto[auto$Renew.Offer.Type=="Offer2",]
Offer12 <-data.frame(rbind(Offer1,Offer2))

ggplot(Offer12, aes(x=Marital.Status, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Marital Status") + ylab("Percent") + 
  ggtitle("Response Rate to Offers 1 & 2 by Marital Status") +
  scale_fill_manual (values = c("darkorange","darkorange4"), name = "Response") + Mytheme

ggplot(Offer12, aes(x=EmploymentStatus, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Employment Status") + ylab("Percent") + 
  ggtitle("Response Rate to Offers 1 & 2 by Employment Status") +
  scale_fill_brewer (palette = "Dark2", name = "Response") + Mytheme

ggplot(Offer12, aes(x=Location.Code, fill = Response)) + 
  geom_bar(stat = "count", position = "fill") +
    xlab("Region") + ylab("Percent") + 
  ggtitle("Response Rate to Offers 1 & 2 by Region") +
  scale_fill_manual (values = c("tan1","tan4"), name = "Response") + Mytheme

```

```{r}
ggplot(auto, aes(x=Monthly.Premium.Auto, y=Customer.Lifetime.Value)) + 
  geom_point(aes(col = as.factor(Number.of.Open.Complaints), size = Months.Since.Last.Claim)) +
  scale_color_brewer(palette="Spectral", name = "Number of Open Complaints") + xlab("Monthly Premium") + 
  ylab("Customer Lifetime Value") + ggtitle("Key Factors to Customer Lifetime Value") +
  labs(size ="Months Since Last Claim") + theme(axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face ="bold"), 
          plot.title = element_text(size = 16, face = "bold"))
```


```{r}
ggplot(auto, aes(x=Total.Claim.Amount, y=Customer.Lifetime.Value)) + 
  geom_point(aes(col = as.factor(Number.of.Policies), size = Months.Since.Policy.Inception)) +
  scale_color_brewer(palette="Spectral", name = "Number of Policies") + xlab("Total Claim Amount") + 
  ylab("Customer Lifetime Value") + ggtitle("Key Factors to Customer Lifetime Value") +
  labs(size ="Months Since Policy Inception") + theme(axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face ="bold"), 
          plot.title = element_text(size = 16, face = "bold"))
```

```{r}
autoCorr <- auto[,c(3,10,13:17,22)]
library(corrplot)
library(corrgram)
colnames(autoCorr) <- c("CLV", "Income", "Mon. Premium", "Mon Since Claim", "Mons Since Inception",
                        "Open Complaints", "Num of Poicies", "Total Claim Amt.")
autoCorr <- cor(autoCorr)
corrplot(autoCorr, method = "number")
```


```{r}
hist(auto$Customer.Lifetime.Value)
summary(auto$Customer.Lifetime.Value)
sd(auto$Customer.Lifetime.Value)
```
Customer Lifetime Values (CLV) at FMTV Insurance ranges from 1,898 to 83,325 with a median of 5,780 and a standard deviation of 6,871. The mean CLV of 8,005 is greater than the median indicating a right skewed distribution with a decreasing number of customers as CLV increases. 

```{r}
library(RColorBrewer)
library(ggplot2)
CLV_PolicyType <- ggplot(auto, aes(x=Policy.Type, y=Customer.Lifetime.Value, fill = Policy ))+
        geom_col() + xlab("Policy Type") + ylab("Customer Lifetime Value") +
        ggtitle("Customer Lifetime Value by Policy Type and Policy") +
         scale_fill_brewer(palette = "Set1")
CLV_PolicyType
```
Customers who purchase Personal Auto insurance provide the lion's share of customer lifetime value to FMTV insurance.

```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)
CLV_Vehicle <- ggplot(auto, aes(x=Vehicle.Size, y=Customer.Lifetime.Value, 
              fill = Vehicle.Class ))+geom_col() + 
        xlab("Vehicle Type") + ylab("Customer Lifetime Value") + 
        ggtitle("Customer Lifetime Value by Vehicle Type and Vehicle Class") +
        scale_fill_brewer(palette = "Dark2")
        
CLV_Vehicle
```
Customers with Midsize vehicles offer the highest customer lifetime values with Four-Door and SUV models

```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)
CLV_Region <- ggplot(auto, aes(x=State, y=Customer.Lifetime.Value, 
              fill = Location.Code ))+geom_col() + 
        xlab("State") + ylab("Customer Lifetime Value") +
        ggtitle("Customer Lifetime Value by State and Region") +
        scale_fill_brewer(palette = "Set1")
        
CLV_Region
```
California and Oregon are the states with the highest Customer Lifetime Value (CLV) customers. Suburban areas have higher CLV customers than other location codes.

```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)
CLV_Marital <- ggplot(auto, aes(x=Marital.Status, y=Customer.Lifetime.Value, 
              fill = Gender ))+geom_col() + 
        xlab("Marital") + ylab("Customer Lifetime Value") +
        ggtitle("Customer Lifetime Value by Marital Status and Gender") +
        scale_fill_brewer(palette = "Set1")
        
CLV_Marital
```



```{r}
library(ggplot2)
ggplot(auto, aes(x=State, y=Customer.Lifetime.Value, fill = Response)) + 
  geom_col(position = "fill")
```

```{r}
library(RColorBrewer)
library(ggplot2)
ggplot(auto, aes(x=Renew.Offer.Type, y=Customer.Lifetime.Value, fill = Response)) + geom_col() +
       xlab("Renewal Offer Type") + ylab("Customer Lifetime Value") +
      ggtitle("Customer Lifetime Value by Renewal Offer and Response") +
      scale_fill_brewer(palette = "Set1")
```

**CLV and State Distribution**
```{r}
CLV.statedf <- data.frame(auto$State, auto$Customer.Lifetime.Value)
CLV.state <- tapply(CLV.statedf$auto.Customer.Lifetime.Value, CLV.statedf$auto.State, sum)
CLV.state
plot.CLV.state <- barplot(CLV.state, main = "CLV by State")
plot.CLV.state
```

**CLV mean by state**
```{r}
CLV.stateMean <- tapply(CLV.statedf$auto.Customer.Lifetime.Value, CLV.statedf$auto.State, mean)
CLV.stateMean
plot.CLV.stateMean <- barplot(CLV.stateMean, main = "CLV Mean By State")
plot.CLV.stateMean
```

**CLV and Education Distribution**
```{r}
CLV.educationdf <- data.frame(auto$Customer.Lifetime.Value, auto$Education)
CLV.education <- tapply(CLV.educationdf$auto.Customer.Lifetime.Value, CLV.educationdf$auto.Education, sum)
CLV.education
plot.CLV.education <- barplot(CLV.education, main = "CLV by Education Level", 
                              cex.names = 0.8) +
  Mytheme2
plot.CLV.education
```

**CLV mean by education**
```{r}
CLV.educationMean <- tapply(CLV.educationdf$auto.Customer.Lifetime.Value, CLV.educationdf$auto.Education, mean)
CLV.educationMean
plot.CLV.educationMean <- barplot(CLV.educationMean, main = "CLV vs. Education Distribution")
plot.CLV.educationMean
```

**CLV and Gender**
```{r}
CLV.genderdf <- data.frame(auto$Customer.Lifetime.Value, auto$Gender)
CLV.gender <- tapply(CLV.genderdf$auto.Customer.Lifetime.Value, CLV.genderdf$auto.Gender, mean)
CLV.gender
plot.CLV.gender <- barplot(CLV.gender, main = "Mean CLV by Gender")
plot.CLV.gender
```

**CLV and Incomebin**
```{r}
CLV.IBdf <- data.frame(auto$Customer.Lifetime.Value, auto$IncomeBin)
CLV.IB <- tapply(CLV.IBdf$auto.Customer.Lifetime.Value, CLV.IBdf$auto.IncomeBin, mean)
CLV.IB
plot.CLV.IB <- barplot(CLV.IB, main = "Mean CLV by Income", las = 2, cex.names = 0.65)
plot.CLV.IB
```

## Discover the correlations between variables
```{r}
premium.CLV <- ggplot(data=auto, aes(x=Customer.Lifetime.Value, y=Monthly.Premium.Auto)) +geom_point()
premium.CLV <- premium.CLV + ggtitle("Potential Correlation Between CLV and Monthly Premium Auto")
premium.CLV 
# The positive correlation is observed. 
```

**Potential Correlation Between CLV and Number Of Open Complaints**
```{r}
complaints.CLV <- ggplot(data=auto, aes(x=Customer.Lifetime.Value, y=Number.of.Open.Complaints)) +geom_point()
complaints.CLV <- complaints.CLV + ggtitle("Potential Correlation Between CLV and Number Of Open Complaints")
complaints.CLV
# More complaints tend to receive lower CLV
```

**Total complaints**
```{r}
totalClaim.CLV <-  ggplot(data=auto, aes(x=Customer.Lifetime.Value, y=Total.Claim.Amount)) +geom_point()
totalClaim.CLV <- totalClaim.CLV + ggtitle("Potential Correlation Between CLV and Total Complaints")
totalClaim.CLV
ggplot(data=auto, aes(x=Customer.Lifetime.Value, y=Total.Claim.Amount)) +geom_point()
```

**Coverage VS CLV**
```{r}
coverage.CLV <- ggplot(data = auto, aes(x = Customer.Lifetime.Value, y = Coverage)) + geom_point()
coverage.CLV <- coverage.CLV + ggtitle("Potential Correlation Between CLV and Coverage")
coverage.CLV
# Interestinly, the coverage does not affect much about CLV
```

**Employment VS CLV**
```{r}
employeement.CLV <- ggplot(data = auto, aes(x = Customer.Lifetime.Value, y = EmploymentStatus)) + geom_point()
employeement.CLV <- employeement.CLV + ggtitle("Potential Correlation Between CLV and Employement Status")
employeement.CLV
```
i
**Sales Channel vs CLV**
```{r}
SalesChannel.CLV <- ggplot(data = auto, aes(x = Customer.Lifetime.Value, y = Sales.Channel)) + geom_point()
SalesChannel.CLV <- SalesChannel.CLV + ggtitle("Potential Correlation Between CLV and Sales Channel")
SalesChannel.CLV
```

**Vechicle vs CLV**
```{r}
vehicleClass.CLV <- ggplot(data = auto, aes(x = Customer.Lifetime.Value, y = Vehicle.Class)) + geom_point()
vehicleClass.CLV <- vehicleClass.CLV + ggtitle("Potential Correlation Between CLV and Vehicle Class")
vehicleClass.CLV
```

**Vehicle Size vs CLV**
```{r}
Vehicle.Size.CLV <- ggplot(data = auto, aes(x = Customer.Lifetime.Value, y = Vehicle.Size)) + geom_point()
Vehicle.Size.CLV <- Vehicle.Size.CLV + ggtitle("Potential Correlation Between CLV and Vehicle Size")
Vehicle.Size.CLV
```
